Using Metric Space Indexing for Complete and Efficient Record Linkage

### 2017-11-21 paper.tex
Update on Overleaf.

Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair of records being classified as a
link or non-link depending on the degree of similarity between them.
Record linkage is usually performed in a three-step process: first
groups of similar candidate records are identified using indexing; 
pairs within the same group are then compared in more detail, and
finally classified.
%
Even state-of-the-art indexing techniques, such as Locality Sensitive
Hashing, have potential drawbacks. They may fail to group together some
true matching records with high similarity. Conversely, they may group
records with low similarity, leading to high computational overhead.
%
We propose using metric space indexing to perform \emph{complete} record
linkage, which results in a parameterless record linkage process
combining indexing, comparison and classification into a single step
delivering complete and efficient record linkage.

@@

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-Tree. The brute-force approach is used to establish the
threshold value that yields the best possible linkage for the Cora
dataset. The threshold value is varied between 0 and 250,
\Cref{fig-cora-brute} presents the precision, recall, and F-measure
values. The linkage results have high precision and low recall for low
threshold values, and high recall and low precision for high threshold
values. For this dataset, the maximum F-measure value is achieved using
a threshold value of 70. This value is data dependent; for different
datasets the maximum F-measure will correspond to different thresholds.
In the rest of this section we fix the threshold value to be 70 which
maximises the F-measure for this dataset.

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
\caption{Cora - Brute Force}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
\caption{Cora - M-Tree}
\end{subfigure}

@@

\caption{Linkage results on the Cora data set}
\label{demography-quality-mtree}
\end{figure}

@@

\Cref{comparison-of-results-cora} shows the results of the Cora
experiments. The table shows the parameters for the experiment, the
number of distance comparisons made, the precision, recall and the F1
measure achieved by each algorithm.  In the \emph{Linker} column the algorithm name is followed by its
parameters: for LSH the number of the bands followed by the band size, and for traditional blocking the fields used
for blocking.

M-tree yields the same linkage quality as brute-force, although using a significantly lower
number of comparisons.

@@

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\caption{F-Measure values of all LSH variations vs M-Tree for the Skye data set}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\caption{F-Measure values of all LSH variations vs M-Tree for the Kilmarnock data set}
\end{subfigure}

\caption{Linkage results on the demographic data set}
\label{demography-quality-mtree}
\end{figure}

% \begin{figure}
% \includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
% \caption{Comparison of quality attainable on Skye dataset using M-tree and various LSH configurations\label{mtree-lsh-quality}}
% \end{figure}

@@

However, this paper also highlights some unexpected results. Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase. A second unexpected result is that shown in \Cref{fig-cora-brute}, namely that LSH can yield higher precision than is achieved by a complete method such as M-Tree as was discussed in Sect.~\ref{sec-approach}.

>>>>

Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair of records being classified as a
link or non-link depending on the degree of similarity between them.
Record linkage is usually performed in a three-step process: first
groups of similar candidate records are identified using indexing; 
pairs within the same group are then compared in more detail, and
finally classified.
%
Even state-of-the-art indexing techniques, such as Locality Sensitive
Hashing, have potential drawbacks. They may fail to group together some
true matching records with high similarity. Conversely, they may group
records with low similarity, leading to high computational overhead.
%
We propose using metric space indexing to perform \emph{complete} record
linkage, which results in a parameter-free record linkage process
combining indexing, comparison and classification into a single step
delivering complete and efficient record linkage.

@@

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree. The brute-force approach is used to establish the
threshold value that yields the best possible linkage for the Cora
dataset. The threshold value is varied between 0 and 250,
\Cref{cora-quality} presents the precision, recall, and F-measure
values. The linkage results have high precision and low recall for low
threshold values, and high recall and low precision for high threshold
values. For this dataset, the maximum F-measure value is achieved using
a threshold value of 70. This value is data dependent; for different
datasets the maximum F-measure will correspond to different thresholds.
In the rest of this section we fix the threshold value to be 70 which
maximises the F-measure for this dataset.

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
\caption{Cora - Brute Force}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
\caption{Cora - M-tree}
\end{subfigure}

@@

\caption{Linkage results on the Cora data set}
\label{cora-quality}
\end{figure}

@@

\Cref{comparison-of-results-cora} shows greater detail for a subset of
the Cora experiments, with the distance threshold fixed at 70. The table
shows the parameters for the experiment, the number of distance
comparisons made, the precision, recall and the F1 measure achieved by
each algorithm.  In the \emph{Linker} column the algorithm name is
followed by its parameters: for LSH the number of the bands followed by
the band size, and for traditional blocking the fields used for
blocking. The number of distance comparisons is reported as a
machine-independent proxy for execution cost, since code profiling shows
that distance calculations are dominant.

M-tree yields the same linkage quality as brute-force, although using a significantly lower
number of comparisons. This is as expected, since both techniques are complete.

Three of the LSH configurations give similar quality

@@

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\caption{Linkage quality on Skye dataset using M-tree and all LSH configurations}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\caption{Linkage quality on Kilmarnock dataset using M-tree and all LSH configurations}
\end{subfigure}

\caption{Linkage results on the demographic datasets}
\label{demography-quality-mtree}
\end{figure}

@@

However, this paper also highlights some unexpected results. Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase. A second unexpected result is that shown in \Cref{quality-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.
